\LARGE Micro- versus meso-scale population models for ecological forecasting under climate change

### 2014-11-25 manuscript/ms_MicroMeso.Rnw
updated model selection section

\noindent where $x$ is the log of genet size (or log of quadrat areal cover), $\beta_{0,t}$ is a year-specific intercept, $\beta_{s,t}$ is the year-specific slope parameter for size, $\beta_{Q}$ is the random effect of quadrat group location, $\theta$ is a matrix of $i$ climate effects over $t$ years, $\beta_{c,i}$ is the fixed parameter for the effect of the $i$th climate covariate, and $\varepsilon_{t}$ is the error term.

@@

\subsection{Selecting climate covariates}
For both the individual-level IPM (IPM) and the quadrat-based IBM (QBM) we followed the same model selection approach for including climate covariates in vital rate regressions.  Working within the random effects structure described previously, we fit a set of models with all possible combinations of four climate covariates: fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year. The combinations of climate covariates we fit included same-year interactions between precipitation and temperature (e.g., a ppt1$\times$TmeanSpr1 interaction). Finally, we also fit a model with no climate covariates included as fixed effects. In total, this resulted in a set of 25 candidate models.

We based model selection on the Deviance Information Criteria (DIC). However, for many species there were several models with $\Delta$DIC < 4, indicating that each model received generally the same amount of support. Given this ambiguity, we opted to simply select the candidate model with lowest DIC, regardless of the statistical nearness of other models, as the final model to use for simulating the population dynamics. In some cases this resulted in selecting a model with no climate covariates. We had to choose a single statistical model for each vital rate for our population models, but when comparing climate covariate selection between the individual and quadrat levels we acknowledge climate covariates selected in all models with $\Delta$DIC < 4.

\subsection{Computing}
\subsubsection{Model fitting}
All of our analyses (model selection, fitting, and simulating) were conducted in R. We used the analytical Bayesian software INLA to fit candidate vital rate models at the individual and quadrat levels for computational efficiency. This is the case except for the recruitment model at the individual level which required a more complex hierarchical approach. In that case we used the MCMC sampler in JAGS to estimate the posterior distributions of model parameters. After model selection, we used MCMC samplers, again in JAGS, to estimate the posterior distributions of model parameters for all vital rate models at the individual and quadrat levels. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 20,000 iterations, after discarding the initial 10,000 iterations. We assessed convergence visually and using the Gelman diagnostic in the R package `coda.' Scale reduction factors for all parameters were less than 1.02, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each chain for all parameters to be used as randomly drawn values during population simulation.

>>>>

\noindent where $x$ is the log of genet size (or log of quadrat areal cover), $\beta_{0,t}$ is a year-specific intercept, $\beta_{s,t}$ is the year-specific slope parameter for size, $\beta_{Q}$ is the random effect of quadrat group location, $\theta$ is a matrix of $i$ climate effects over $t$ years, $\beta_{c,i}$ is the fixed parameter for the effect of the $i$th climate covariate, and $\varepsilon_{t}$ is the error term. At the quadrat level we did not fit random year effects for survival because those parameters did not reach convergence (see "Computing").

@@

\subsection{Including climate covariates (and avoiding model selection)}
For both the individual-level IPM (IPM) and the quadrat-based IBM (QBM) we included four climate covariates in each vital rate regression: fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year. We included climate covariates as additive effects. We did not consider interactions among climate covariates or among climate covariates and plant size or proportional cover. Climate effects ($\beta_{c}$s) were modeled hierarchically in each vital rate regression so that each species specific climate effect is drawn from an interspecific climate effect distribution. This was necessary to achieve convergence at the quadrat level and we used the same approach at the individual level for uniformity.

\subsection{Computing}
\subsubsection{Model fitting}
All of our analyses (model fitting and simulating) were conducted in Program R. We used the MCMC sampler in JAGS to estimate the posterior distributions of model parameters. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 50,000 iterations, after discarding the initial 20,000 iterations. We assessed convergence visually and using the Gelman diagnostic in the R package `coda.' Scale reduction factors for all parameters were less than 1.02, indicating convergence. For the purposes of introducing stochasticity in our population models, we saved the final 1,000 iterations from each chain for all parameters to be used as randomly drawn values during population simulation.

### 2014-11-25 manuscript/ms_MicroMeso.tex
updated model selection section

\subsection{Selecting climate covariates}
For both the individual-level IPM (IPM) and the quadrat-based IBM (QBM) we followed the same model selection approach for including climate covariates in vital rate regressions.  Working within the random effects structure described previously, we fit a set of models with all possible combinations of four climate covariates: fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year. The combinations of climate covariates we fit included same-year interactions between precipitation and temperature (e.g., a ppt1$\times$TmeanSpr1 interaction). Finally, we also fit a model with no climate covariates included as fixed effects. In total, this resulted in a set of 25 candidate models.

We based model selection on the Deviance Information Criteria (DIC). However, for many species there were several models with $\Delta$DIC < 4, indicating that each model received generally the same amount of support. Given this ambiguity, we opted to simply select the candidate model with lowest DIC, regardless of the statistical nearness of other models, as the final model to use for simulating the population dynamics. In some cases this resulted in selecting a model with no climate covariates. We had to choose a single statistical model for each vital rate for our population models, but when comparing climate covariate selection between the individual and quadrat levels we acknowledge climate covariates selected in all models with $\Delta$DIC < 4.

@@

% latex table generated in R 3.1.1 by xtable 1.7-3 package
% Tue Nov 18 09:15:14 2014
\begin{table}[ht]
\centering
\caption{The proportion of interannual variability in  vital rates explained by the climate covariates. The contribution for growth is defined as: (Climate model - Constant Model)/(Full model - Constant model). The contribution for survival and colonization, where we could not estimate a full model with year random effects, is defined as: (Constant Model - Climate Model)/Constant Model. The values are model deviance estimated from the `coda' package in R.} 
\begin{tabularx}{\textwidth}{llllX}
  \hline
Vital Rate & Constant model & Climate model & Full model & Contribution of climate covariates \\ 
  \hline
Growth & -5566.24 & -5735.32 & -6167.98 & 0.28 \\ 
  Survival & 171.55 & 170.37 &  & 0.01 \\ 
  Colonization & 203.98 & 193.65 &  & 0.05 \\ 
   \hline
\end{tabularx}
\end{table}
% # <<echo=FALSE, fig=TRUE,height=3,width=8.5 >>=
% # library(ggplot2)
% # plotD <- data.frame(Species = rep(growthSSRTable$Species, 3),
% #                     Model = rep(c("Growth", "Survival", "Colonization"), each=4),
% #                     Contribution = c(growthSSRTable[,6],
% #                                      survivalSSRTable[,6],
% #                                      colonizationSSRTable[,6]))
% # ggplot(plotD, aes(x=Species, y=Contribution))+
% #   geom_bar(stat="identity")+
% #   facet_grid(.~Model)
% # @

>>>>

\subsection{Including climate covariates (and avoiding model selection)}
For both the individual-level IPM (IPM) and the quadrat-based IBM (QBM) we included four climate covariates in each vital rate regression: fall through spring precipitation at \emph{t}-1 and \emph{t}-2 (ppt1 and ppt2, respectively) and mean spring temperature at \emph{t}-1 and \emph{t}-2 (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t} is the observation year. We included climate covariates as additive effects. We did not consider interactions among climate covariates or among climate covariates and plant size or proportional cover.

@@

% latex table generated in R 3.1.0 by xtable 1.7-3 package
% Tue Nov 25 11:51:40 2014
\begin{table}[ht]
\centering
\caption{The proportion of interannual variability in  vital rates explained by the climate covariates. The contribution for growth is defined as: (Climate model - Constant Model)/(Full model - Constant model). The contribution for survival and colonization, where we could not estimate a full model with year random effects, is defined as: (Constant Model - Climate Model)/Constant Model. The values are model deviance estimated from the `coda' package in R.} 
\begin{tabularx}{\textwidth}{llllX}
  \hline
Vital Rate & Constant model & Climate model & Full model & Contribution of climate covariates \\ 
  \hline
Growth & -5566.24 & -5735.32 & -6167.98 & 0.28 \\ 
  Survival & 171.55 & 170.37 &  & 0.01 \\ 
  Colonization & 203.98 & 193.65 &  & 0.05 \\ 
   \hline
\end{tabularx}
\end{table}
% # <<echo=FALSE, fig=TRUE,height=3,width=8.5 >>=
% # library(ggplot2)
% # plotD <- data.frame(Species = rep(growthSSRTable$Species, 3),
% #                     Model = rep(c("Growth", "Survival", "Colonization"), each=4),
% #                     Contribution = c(growthSSRTable[,6],
% #                                      survivalSSRTable[,6],
% #                                      colonizationSSRTable[,6]))
% # ggplot(plotD, aes(x=Species, y=Contribution))+
% #   geom_bar(stat="identity")+
% #   facet_grid(.~Model)
% # @
